Vagueness \& Rationality

### 2018-09-03 springer/current/related-work.tex
Update related-work.tex

Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers. Two recent studies can illustrate both issues.

>>>>

A number of answers to Litman’s question have been proposed, but as we have argued elsewhere [REF KvD JPL and REF book], some of the most often cited answers do not hold up to scrutiny. For example,
%
\begin{itemize}
\item It has been proposed that the existence of vague words makes a language more efficient [REF Barwise Cooper]. The idea is that vagueness allows the use of one and the same word (e.g., ``big”) in different situations, which require different standards. For instance, the vague word ``’big” can denote a mouse a well as an elephant. — The idea is plausible, but the problem is that ``big" is efficient not because it is vague (i.e., not because it allows borderline cases) but because it is context dependent. To see this, it suffices to realise that superlatives are not vague, but they {\em are} context dependent: to be ``the biggest $x$" ascribes different sizes to the referent depending on the word $x$. It is for this reason that we can use the word ``biggest” to talk about the biggest mouse, the biggest elephant, and so on. Context-dependence, however, does not imply vagueness.
%
\item It has been proposed that the ability to combine a statement of quantity (along some dimension) with an evaluative statement [REF Veltman]. The idea is that when we say “The patient has a fever”, we do not merely assert that her temperature is above a certain point in the temperature dimension, we also imply that the deviation is large enough to be clinically significant (i.e., something’s not enirely right with the patient). — The problem with this equally plausible idea is that evaluation does not imply vagueness. For example, the medical term “obese” is evaluative (i.e., it’s not healthy to be obese), yet its standard medical definition in terms of Body Mass Index is perfectly crisp.
\end{itemize}
%
The findings on which we are reporting in the present paper will follow a similar pattern: we will explore a possible answer to Litman’s question, only to conclude that it does not stand up to experimental scrutiny.

Charting the utility of vagueness is also the attested aim of a small number of experimental studies. However, to the best of our knowledge, few of these studies have truly focussed on vagueness in the sense on which we are focussing (i.e., they did not address Litman's challenge). Two recent studies can illustrate both issues.
